Research Notes

### 2019-02-18 writeups/notes.tex
correct mask, justification

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)
            + \frac{1}{2k_zH}} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)
            + \frac{1}{2k_zH}} & z < z_0\\
    \end{cases}.
\end{equation}
The extra phase shift comes from $\pd{}{z} \approx ik_z + \frac{1}{2H} \approx
k_z\p*{e^{-\frac{1}{2k_zH}}}$.

@@

>>>>

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)
            + \frac{1}{2k_zH}} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)
            + \frac{1}{2k_zH}} & z < z_0\\
    \end{cases}.\label{eq:anal_forced}
\end{equation}
The extra phase shift comes from $\pd{}{z} \approx ik_z + \frac{1}{2H} \approx
k_z\p*{e^{-\frac{1}{2k_zH}}}$.

@@

\section{Masking Advective Terms}

Recall \autoref{eq:anal_forced} tells us the amplitude of the excited wave in
terms of the forcing function. However, numerically, it appears that the forcing
zone evolves a local mean flow $\bar{U}_0 < 0$ due to the nonlinear terms. The
effect of such a term is to replace $\omega_0 \to \omega_0 - k_x\bar{U}_0 >
\omega_0$ in \autoref{eq:anal_forced}, which has the effect of attenuating the
forcing amplitude by $\sim \frac{1}{1 + \frac{\bar{U}_0}{c_{ph, x}}} \approx
\p*{1 - \frac{\bar{U}_0}{c_{ph, x}}}$.

The cause of this term comes from the viscous terms, which $\sim \partial_t -
\nu \nabla^2$ (our particular implementation differs slightly). In the
$\delta$-function forcing limit, $u_x, u_z$ are only $\mathcal{C}^0$ at $z_0$.
The derivative jumps at $z_0$ with height dictated by the forcing strength.
Thus, $\nabla^2 u_x, \nabla^2 u_z \sim \delta(z - z_0)$, and there is a strong
extra contribution to the EOM at $z_0$.
